---
title: "Sangres Tree Stats and Graphs"
author: "Jonathan Grabowski"
date: "2024-10-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# load packages, read in the data, and do some stuff
```{r}
library(ggplot2)
library(reshape2)
library(readxl)
library(dplyr)
library(gridExtra)
library(tinytex)

SFS4 <- read_excel("data/sfs4 bible 2024.xlsx")
BTN4 <- read_excel("data/btn4 revisit 2024.xlsx")
BTN4dbh <- read_excel("data/btn4 access for C.xlsx")
SFF1 <- read_excel("data/sff1 establishment 2024.xlsx")
SFF2 <- read_excel("data/sff2 establishment 2024.xlsx")
SFF3 <- read_excel("data/sff3 establishment 2024.xlsx")
SFF4 <- read_excel("data/sff4 establishment 2024.xlsx")
SFF5 <- read_excel("data/sff5 establishment 2024.xlsx")
SFF6 <- read_excel("data/sff6 establishment 2024.xlsx")
SFF7 <- read_excel("data/sff7 establishment 2024.xlsx")
SFF8 <- read_excel("data/sff8 establishment 2024.xlsx")
SFF9 <- read_excel("data/sff9 establishment 2024.xlsx")
SFF10 <- read_excel("data/sff10 establishment 2024.xlsx")

#Add plot numbers to treeIDs

SFS4$Tree_num = paste0('SFS4-', SFS4$Tree_num)
BTN4$Tree_num = paste0('BTN4-', BTN4$Tree_num)
SFF1$Tree_num = paste0('SFF1-', SFF1$Tree_num)
SFF2$Tree_num = paste0('SFF2-', SFF2$Tree_num)
SFF3$Tree_num = paste0('SFF3-', SFF3$Tree_num)
SFF4$Tree_num = paste0('SFF4-', SFF4$Tree_num)
SFF5$Tree_num = paste0('SFF5-', SFF5$Tree_num)
SFF6$Tree_num = paste0('SFF6-', SFF6$Tree_num)
SFF7$Tree_num = paste0('SFF7-', SFF7$Tree_num)
SFF8$Tree_num = paste0('SFF8-', SFF8$Tree_num)
SFF9$Tree_num = paste0('SFF9-', SFF9$Tree_num)
SFF10$Tree_num = paste0('SFF10-', SFF10$Tree_num)

# join of BTN4 DBH & Species files
BTN4dbh$Tree_num <- BTN4dbh$TreeID
BTN4 <- left_join(BTN4, BTN4dbh, by = "Tree_num")
BTN4$Species <- BTN4$SpeciesID
BTN4$Notes<-BTN4$Notes.x

#Add plot number to data
SFS4$PlotName <- "SFS4"
BTN4$PlotName <- "BTN4"
SFF1$PlotName <- "SFF1"
SFF2$PlotName <- "SFF2"
SFF3$PlotName <- "SFF3"
SFF4$PlotName <- "SFF4"
SFF5$PlotName <- "SFF5"
SFF6$PlotName <- "SFF6"
SFF7$PlotName <- "SFF7"
SFF8$PlotName <- "SFF8"
SFF9$PlotName <- "SFF9"
SFF10$PlotName <- "SFF10"

#Add treatment status

SFS4$TreatmentStatus <- "Treated"
BTN4$TreatmentStatus <- "Untreated"
SFF1$TreatmentStatus <- "Treated"
SFF2$TreatmentStatus <- "Untreated"
SFF3$TreatmentStatus <- "Untreated"
SFF4$TreatmentStatus <- "Untreated"
SFF5$TreatmentStatus <- "Treated"
SFF6$TreatmentStatus <- "Untreated"
SFF7$TreatmentStatus <- "Treated"
SFF8$TreatmentStatus <- "Treated"
SFF9$TreatmentStatus <- "Untreated"
SFF10$TreatmentStatus <- "Treated"

# Add plot size

SFS4$PlotSize <- 1
BTN4$PlotSize <- 1
SFF1$PlotSize <- 0.25
SFF2$PlotSize <- 1
SFF3$PlotSize <- 0.25
SFF4$PlotSize <- 0.25
SFF5$PlotSize <- 0.25
SFF6$PlotSize <- 0.25
SFF7$PlotSize <- 0.25
SFF8$PlotSize <- 1
SFF9$PlotSize <- 0.25
SFF10$PlotSize <- 0.25
  
#Select columns of interest (make loop at some point)

SFS4 <- SFS4 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG, Notes)
BTN4 <- BTN4 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG, Notes)
SFF1 <- SFF1 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG, Notes)
SFF2 <- SFF2 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG, Notes)
SFF3 <- SFF3 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG, Notes)
SFF4 <- SFF4 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG, Notes)
SFF5 <- SFF5 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG,  Notes)
SFF6 <- SFF6 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG, Notes)
SFF7 <- SFF7 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG, Notes)
SFF8 <- SFF8 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG, Notes)
SFF9 <- SFF9 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG,  Notes)
SFF10 <- SFF10 %>%
  select(PlotName, PlotSize, TreatmentStatus, Tree_num, Species, Condition, DBH, OG, Notes)


# merge all the plots into one dataframe
SFS4$Condition <- as.numeric(SFS4$Condition)
merged_plots <- bind_rows(SFS4, BTN4, SFF1, SFF2, SFF3, SFF4, SFF5, SFF6, SFF7, SFF8, SFF9, SFF10)

# add a column to define MOG as Y or NA
merged_plots <- merged_plots %>%
  mutate(MOG = case_when(OG == "Y" | DBH > 30 ~ "Y"))
             
# calculate tree counts
merged_summary <- merged_plots %>%
  group_by(PlotName, PlotSize, TreatmentStatus) %>%
  summarize(Tree_count = n(), meanDBH = mean(DBH, na.rm = TRUE))

# calculate MOG stats
merged_MOG <- merged_plots %>%
  filter(MOG == "Y")%>%
  group_by(PlotName, PlotSize, TreatmentStatus)%>%
  summarize(MOG_count = n())

#calculate live MOG stats
merged_MOG_live <- merged_plots %>%
  filter(MOG == "Y" & Condition != 5 & Condition != 2) %>%
  group_by(PlotName, PlotSize, TreatmentStatus) %>%
  summarize(MOG_live_count = n())

#calculate OG stats
merged_OG <- merged_plots %>%
  filter(OG == "Y") %>%
  group_by(PlotName, PlotSize, TreatmentStatus) %>%
  summarize(OG_count = n())

# merge them all together
merged_summary$MOG_count <- merged_MOG$MOG_count
merged_summary$MOG_live_count <- merged_MOG_live$MOG_live_count
merged_summary$OG_count <- merged_OG$OG_count

# calculate TPH
merged_summary$TPH <- merged_summary$Tree_count/merged_summary$PlotSize
est_data<-merged_summary

# Add tree colors 

add_tree_colors <- function(df){
  df <- df%>%
    mutate(color=
             case_when(Species == "ABCO" ~ "#4EDFC7",
                       Species == "ACGL" ~ "#C21E56",
                       Species == "JUMO" ~ "#FFC0CB",
                       Species == "JUSC" ~ "#95658B",
                       Species == "PIED" ~ "#FFD700",
                       Species == "PIPO" ~ "#2E8B57",
                       Species == "PIST" ~ "#89CFF0",
                       Species == "PRVI" ~ "#9F2B68",
                       Species == "PSME" ~ "#808080",
                       Species == "QUGA" ~ "#5D3FD3",
                       Species == "QUUN" ~ "#CC5500",
                       Species == "SASC" ~ "#E3963E",
                       Species == "unknown" ~ "#AFE1AF",
                       Species == "NA" ~ "#808080"
                       
             ))
  
}

merged_plots <- add_tree_colors(merged_plots)

# some treatment level stats

treatment_summary <- merged_summary %>%
  group_by(TreatmentStatus)%>%
  summarize(Tree_count = sum(Tree_count), MOG_count = sum(MOG_count), MOG_live_count = sum(MOG_live_count), Surveyed_ha = sum(PlotSize))

```


# Makes a graph of the number of trees sampled per plot
```{r}
ggplot(est_data, aes(x= reorder(PlotName, Tree_count), y= Tree_count)) +
  geom_bar(stat="identity", color = 'black', aes(fill = TreatmentStatus)) +
  xlab("") +
  ylab("Number of trees") +
  ylim(0, 2000) +
  scale_fill_manual(values=c("#00aedb", "#d11141")) +
  theme_classic() +
  theme(legend.title = element_blank())
```


# Makes a graph showing the number of MOG (mature old growth) sampled per plot
```{r}
ggplot(est_data, aes(x= reorder(PlotName, MOG_count), y= MOG_count)) +
  geom_bar(stat="identity", color = 'black', aes(fill = TreatmentStatus)) +
  xlab("") +
  ylab("Number of mature old growth") +
  ylim(0,200) +
  scale_fill_manual(values=c("#00aedb", "#d11141")) +
  theme_classic() +
  theme(legend.title = element_blank())
```


# Makes a graph showing the number of live MOG (mature old growth) sampled per 
# plot
```{r}
ggplot(est_data, aes(x= reorder(PlotName, MOG_live_count), y= MOG_live_count)) +
  geom_bar(stat="identity", color = 'black', aes(fill = TreatmentStatus)) +
  xlab("") +
  ylab("Number of live mature old growth") +
  ylim(0, 200) +
  scale_fill_manual(values=c("#00aedb", "#d11141")) +
  theme_classic() +
  theme(legend.title = element_blank())
```


# Makes a graph showing the number of OG (old growth) sampled per plot
```{r}
ggplot(est_data, aes(x= reorder(PlotName, OG_count), y= OG_count)) +
  geom_bar(stat="identity", color = 'black', aes(fill = TreatmentStatus)) +
  xlab("") +
  ylab("Number of old growth") +
  scale_fill_manual(values=c("#00aedb", "#d11141")) +
  ylim(0,200) +
  theme_classic() +
  theme(legend.title = element_blank())
```


# Makes a graph showing the number of MOG per ha per plot
```{r}
MOGperHectare <- est_data$MOG_count / est_data$PlotSize

est_data$MOGperha <- MOGperHectare 

ggplot(est_data, aes(x= reorder(PlotName, MOGperha), y= MOGperha)) +
  geom_bar(stat="identity", color = 'black', aes(fill = TreatmentStatus)) +
  xlab("") +
  ylab("Mature old growth density (trees per ha)") +
  scale_fill_manual(values=c("#00aedb", "#d11141")) +
  ylim(0,200) +
  theme_classic() +
  theme(legend.title = element_blank())
```

--------------------------------------------------------------------------------

## DATA BY TREATMENT

# Graph of number of trees by treatment
```{r}
TreatedNumTrees <- sum(subset(est_data, TreatmentStatus == 'Treated')$Tree_count)

UntreatedNumTrees <- sum(subset(est_data, TreatmentStatus == 'Untreated')$Tree_count)

NumTrees <- c(TreatedNumTrees, UntreatedNumTrees)

NumTreesByTreatment <- data.frame(Treatment, NumTrees)

ggplot(NumTreesByTreatment, aes(x = Treatment, y = NumTrees)) +
  geom_bar(stat = "identity", color = 'black', aes(fill = Treatment)) +
  xlab("") +
  ylab("Number of trees") +
  ylim(0, 5000) +
  scale_fill_manual(values=c("#00aedb", "#d11141")) +
  theme_classic() +
  theme(legend.position = "none")
```

